% !TEX root = ../Thesis.tex
\chapter{Evaluation}
\label{c:evaluation}

This chapter is separated into 

\section{Goal}
The evaluation has two goals the correctness as well as the impact of data freshness onto different kinds of workloads.

Verify and validate the correctness as well as the completeness of the implementation based on several characteristics.
These include the correct execution of lazy replication, the possibility to refresh statements on demand.

Impact of the replication engine on the underlying performance, if freshness indeed increases the overall parallel writes on the system.
Or if it is just marginally lower than before. Also compare this to the overall introduced overhead. And if the change was wort it

\section{Correctness}

The correctness of the introduced solution mainly focuses on two parts. For one the replication behaviour, to verify if each lazy replication is carried out correctly.
And if not verify that reasonable counter measures are in place and apply them. This is crucial since we do not compare the footprint or the integrity of the data after 
a replication update. Rather we compare on a high level the metadata 
(i.e. if the number of modifications and the commit timestamp after the data replication are equal on primary and secondary node) of two replicas. 

The second part of the validation process focuses on the retrieval of outdated nodes. Although we always have a fall back to the primary placements as described in section \ref{sec:freshness_selection},
we still want to avoid excessive locking to parallelize requests to ultimately speed up the average response time.

\section{Benchmarks}

\todo{Explain why it is necessary to verify the solution with different combinations}

\subsection{Evaluation Environment}
\todo{Elaborate and thoroughly explain why the environment was chosen and how the test was executed for the sake of rerpoducibility}


\subsection{Evaluation Procedure}
The following steps outline the procedure for benchmarking data freshness within Polypheny-DB.

\todo{ Execute benchmarks on multimodal dbs. as well as different kind of }

\todo{ Talk about implementation of freshness characteristic in many query languages}


\subsection{Results}

\section{Discussion}
\label{sec:discussion}
The result generally show